Using Distributed Information Stores for Patient-Clinician Health Data Sharing

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Storing and representation together

\section{Storing activity diaries}

@@

@@

%% In order to do this, the data from each device should be collected and archived, so it can
%% be viewed together in an appropriate interface. The data collected by patients' devices has
%% the potential to create a large volume of data, and therefore particular
%% consideration should be paid about who and where the data will be stored. In particular, it is likely
%% that existing providers of storage for electronic health records will not want to be burdened with the
%% administration costs of storing additional data. The longevity of third-party storage providers is
%% also a concern, because there has been a history of short-notice closures of EHR providers (such as Google Health).

\section{Representing activity diaries}

Particular care should be taken when deciding on a format to store the activity diaries. Ideally
a format will be machine-readable, using open standards (for example XML, RDF or JSON) and
following existing schematic/ontological knowledge modelling conventions. There are a 
number of medical ontologies that already exist, as well as vocabularies and systems (such as
SNOMED CT \cite{stearns2001snomed} and IDC-10) that focus on the medical domain. The hope is that by using
such systems, that the data collected by patients can be analysed more rapidly by clinical systems
and staff.

One option is to embrace the Semantic Web, because it offers several key benefits:

\begin{itemize}

\item Distributed system with universal addresses and standard mature protocols
\item Open extensible ontology system and existing domain ontologies
\item Existing library and tool support and expertise
\item Interoperability with existing systems

\end{itemize}

Even if personal health records or systems do not use the Semantic Web, there are existing
techniques to integrate with their data.

>>>>

\section{Storing and Representing Activity Diaries}

@@

The data collected by patients' devices has the potential to create a large volume of highly structured data. This is in contrast to
EHRs which tend to contain individual records for each clinical session.

To this challenge we bring our contribution of a common ontology for recording and encompassing patient activities, which we have begun to
connect to SNOMED-CT, in order to enable semantic interoperability with existing systems.

@@

%% In order to do this, the data from each device should be collected and archived, so it can
%% be viewed together in an appropriate interface. The data collected by patients' devices has
%% the potential to create a large volume of data, and therefore particular
%% consideration should be paid about who and where the data will be stored. In particular, it is likely
%% that existing providers of storage for electronic health records will not want to be burdened with the
%% administration costs of storing additional data. The longevity of third-party storage providers is
%% also a concern, because there has b een a history of short-notice closures of EHR providers (such as Google Health).

%Particular care should be taken when deciding on a format to store the activity diaries. Ideally
%a format will be machine-readable, using open standards (for example XML, RDF or JSON) and
%following existing schematic/ontological knowledge modelling conventions. There are a 
%number of medical ontologies that already exist, as well as vocabularies and systems (such as
%SNOMED CT \cite{stearns2001snomed} and IDC-10) that focus on the medical domain. The hope is that by using
%such systems, that the data collected by patients can be analysed more rapidly by clinical systems
%and staff.
%
%One option is to embrace the Semantic Web, because it offers several key benefits:
%
%\begin{itemize}
%
%\item Distributed system with universal addresses and standard mature protocols
%\item Open extensible ontology system and existing domain ontologies
%\item Existing library and tool support and expertise
%\item Interoperability with existing systems
%
%\end{itemize}
% 
%Even if personal health records or systems do not use the Semantic Web, there are existing
%techniques to integrate with their data.
